Study of the hadronization of charged pions

### 2017-01-13 1-Answers-Review.tex

P15 bottom, “These cuts minimize the kaon...”

P18, 6 lines from the bottom - “As Figure 14 ...”

P20, L1 of section 2.3.2 “... in figure 15 a few ...”

P21, L2, “... processed by CLAS software (GSIM, GPP and user\_ana) [what do these mean to a 
non-CLAS reader? supply references] to simulate the detector and the reconstruction process in 
a fashion similar to that done with the experimental data.”
Start of next paragraph, “The simulated data are processed like the experimental data by 
applying ...”  “ ... quite well the detector response, yet ...” \\

{\it Not sure what reference we should add here. The analysis note is not only a technical document
it is also an internal document.} \\

@@

P36, second paragraph of section 3.1.1 -
 change conciliate to reconcile

P38, Para 3, L2, change signification to significance

P38, bottom, “However, our results (figure 32(left)) show no dependence besides ... and what is 
left will be washed ...”

P41, first paragraph, what does HERMES say about Q2 dependence?  Line 4, “... in figure 34, 
does not indicate ...” \\

{\it HERMES do see an increase which is not statistically significant 
(see figures 2 and 3 of the note).} \\

@@

P41, bottom paragraph, “... wide A coverage gives a outright indication ... found to be much 
smaller than that seen by HERMES [15], which ....from BDMPS [29] correlate pt\^2 with the 
nuclear radius.”

P42, top paragraph, last line, “colored parton does not interact with the whole nucleus which 
limits the nuclear effect.”

P42 second paragraph, L2, “they predict a rise of delta pt\^2 with Q\^2”
Same paragraph, L4, “... 38) gives a similar result.” \\

{\it Sorry for the clumsy English in some parts. We have implemented all the 
suggested changes and some more. This issue should now be solved.} \\

\end{document}

### 2017-01-13 chap5.tex
Finished English corrections

To identify negatively charged pions, we select negative tracks which didn't verify electrons' identification cuts. Pions are detected in the angular range from $\sim$10 to $\sim$140 degrees using only DC and SC signals.

@@

The identification of positively charged pions is similar to that of negative 
pions. However, the time of flight plot is significantly more busy (see figure 
\ref{PipTOF}), indicating a significant contamination from $K^+$s and protons at a
higher momentum. As the CC is not efficient enough for a hadron separation, the 
numerous kaons and protons should be removed by the following tighter TOF cuts:

@@

These cuts permit to minimize the kaon contamination, below 2.5~GeV/c, and the 
proton contamination for all momenta. The kaon contribution cannot be avoided, however, it should remain small ($\sim$ 3\% according to simulation) 
with only a small effect on our final results (see section \ref{SysId} 
for more details). Because protons could lead to even more contamination than kaons, 
a stricter cut is used at a high momentum. This cut is also justified by 
HERMES data \cite{Airapetian:2007vu}, which showed a very different behavior of protons compared to other hadrons.

@@

To check the quality of the runs\footnote{Runs were formed roughly from two hours
of data, but they were much smaller in case of problems.}, we monitored the ratio of scattered electrons yields obtained from solid and liquid targets. In case of the beam hitting other materials in the beam-line or of ony other problem that would affect our final result, 
this ratio will be off and indicate the problematic runs. A figure 
\ref{DataQ} shows electrons yields' ratios for different nuclear target; we fitted these ratios to extract the mean value for each target, then we discarded runs that were 5~$\sigma$ away from the appropriate mean. Must note that these ratios were consistent with our target's thicknesses given in \cite{Hakobyan:2008zz} except for a carbon target. For that reason, the density of the latter was recently remeasured, and its new value was finally matching our data\footnote{i.e. (1.747+/-0.0007)g/cm$^3$ instead of 2.235 g/cm$^3$}.

@@

To provide some insight into the quality of our data, we present in figure~\ref{fig:prelim} few preliminary results before the application of any corrections. These preliminary results will be used also to illustrate the effects of the corrections discussed below.

@@

The acceptance correction consists of applying weights to the experimental data
event by event to correct for any detection inefficiencies of used detectors.
Incidentally, it also corrects for other small detection issues, such as misidentification and re-scattering on detectors' materials. The quality 
of the correction depends on the ability of the simulation to reproduce the 
experiment, the accumulated statistics, and the size of interfering effects, such 
as the bin migration. For this correction, we applied the method from the 
approved and published analysis from the same eg2 data~\cite{ElFassi:2008}.

@@

To correct for acceptance effects, we simulated a total of 100 million events 
per target ($^2$H, C, Fe and Pb) using the PYTHIA \cite{Sjostrand:2006za} 
event generator that was slightly modified to include Fermi motion effects. The 
generated events are processed by CLAS softwares (GSIM, GPP and user\_ana) 
to simulate the detector and the reconstruction process similarly to the experimental data.
% TODO GPP factors

Then the simulated data are processed in a similar way than the experimental data by 
applying the cuts described in the section \ref{sec:pid}. Overall the 
simulation reproduces quite well the detectors responses, yet two issues might 
affect us and have to be understood. First, the efficiency of the CC is 
overestimated in the simulation. On the electron side the signal is a little stronger in the simulation (11 
photo-electrons) compared to experimental data (8 photo-electrons), but this 
feature should not affect us too much, because we are cutting only the tail of the 
distribution in both cases. Second, since in the simulation the beam is perfectly aligned with CLAS, the vertex cuts do not need to be shifted from one sector to an other. Therefore, we do not apply the shift from table \ref{tab:vertex} to the simulated data; the resulting cuts are shown in figure~\ref{simvertex}.

@@

We used a 4-dimensional binning to divide the large phase space of our two particles' final 
state. To evaluate systematic errors associated with this correction, we compared the
method using two different sets of kinematic variables, as presented in 
table~\ref{tab:AcceptBinning} and \ref{tab:AcceptBins}. The first set of variables is the one actually used for the
analysis and the second set is only used to evaluate the systematic errors, so in the 
following, the formulas are given only for the first set. The total number of bins is 
constrained by the amount of generated data and must be maintained reasonably low in order 
to keep a good statistical precision and a small bin migration.

@@

Because of reconstruction errors or scattering on detectors' material, it 
is possible to associate a particle with the wrong target. To estimate this 
effect, we looked, in the experimental data, at the number of reconstructed 
events upstream and downstream of our targets, where nothing should be 
detected. Thus, we defined two test regions (see figure \ref{fig:targetleak}) 
to evaluate the associated contamination between the targets. The upstream 
region, named test region 1, is of the size of the solid target's selection 
cut to evaluate the contamination from the liquid target. The downstream region, 
named test region 2, covers the same size of the liquid target's selection cut 
to estimate the leak from the solid target. The distance between detection and 
test regions is identical to the distance between the two targets' detection 
regions.

@@

The point-to-point errors include contributions from the bad particle 
identifications and acceptance effects. The normalization error, which is 
independent of kinematic variables, are due to acceptance effects, target 
misidentification and isospin correction; the constributions are summarized 
in table~\ref{tab:sysid}.

>>>>

Pions are detected in the angular range from $\sim$10 to $\sim$140 degrees 
using only DC and SC signals. To identify negatively charged pions, we select 
negative tracks which were not identified as electrons already.

@@

The identification of positively charged pions is similar to that of negative 
pions. However, the time of flight plot is significantly more busy (see figure 
\ref{PipTOF}), due to an important contamination from $K^+$s and protons at a
high momentum. As the CC is not efficient enough for a hadron separation, the 
numerous kaons and protons should be removed by the following tighter TOF cuts:

@@

These cuts minimize the kaon contamination, below 2.5~GeV/c, and the 
proton contamination for all momenta. The kaon contribution cannot be avoided, however, it should remain small ($\sim$ 3\% according to simulation) 
with only a small effect on our final results (see section \ref{SysId} 
for more details). Because protons could lead to even more contamination than kaons, 
a stricter cut is used at a high momentum. This cut is also justified by 
HERMES data \cite{Airapetian:2007vu}, which showed a very different behavior of protons compared to other hadrons.

@@

To check the quality of the runs\footnote{Runs are typically stopped every two hours
of data taking, but they can be smaller when experimental problems arrise.}, we 
monitored the ratio of scattered electrons yields obtained from solid and liquid 
targets. In case of the beam hitting other materials in the beam-line or of any 
other problems that would affect our final result, 
this ratio will be off and indicate the problematic runs. In figure 
\ref{DataQ}, we show the electron yield ratios for different nuclear target; we 
fitted these ratios to determine the mean value for each target, then we discarded 
runs that were 5~$\sigma$ away from the appropriate mean. Must note that these 
ratios were consistent with our target's thicknesses given in \cite{Hakobyan:2008zz} 
except for a carbon target. For that reason, the density of the latter was 
recently remeasured, and its new value was finally matching our 
data\footnote{i.e. (1.747+/-0.0007)g/cm$^3$ instead of 2.235 g/cm$^3$}.

@@

To provide some insight into the quality of our data, we present in 
figure~\ref{fig:prelim} a few preliminary results before the application of 
any corrections. These preliminary results will be used also to illustrate 
the effects of the corrections discussed below.

@@

The acceptance correction consists of applying weights to the experimental data
event by event to correct for any detection inefficiencies of used detectors.
Incidentally, it also corrects for other small detection issues, such as 
misidentification and re-scattering on detector material. The quality 
of the correction depends on the ability of the simulation to reproduce the 
experiment, the accumulated statistics, and the size of interfering effects, such 
as the bin migration. For this correction, we applied the method from the 
approved and published analysis from the same eg2 data~\cite{ElFassi:2008}.

@@

To correct for acceptance effects, we simulated a total of 100 million events 
per target ($^2$H, C, Fe and Pb) using the PYTHIA \cite{Sjostrand:2006za} 
event generator that was slightly modified to include Fermi motion effects. The 
generated events are processed by CLAS softwares (GSIM, GPP and user\_ana) 
to simulate the detector and the reconstruction process in a fashion similar 
to that done with the experimental data.
% TODO GPP factors

The simulated data are processed like the experimental data by 
applying the cuts described in the section \ref{sec:pid}. Overall the 
simulation reproduces quite well the detector response, yet two issues might 
affect us and have to be understood. First, the efficiency of the CC is 
overestimated in the simulation. On the electron side the signal is a little stronger in the simulation (11 
photo-electrons) compared to experimental data (8 photo-electrons), but this 
feature should not affect us too much, because we are cutting only the tail of the 
distribution in both cases. Second, since in the simulation the beam is perfectly aligned with CLAS, the vertex cuts do not need to be shifted from one sector to an other. Therefore, we do not apply the shift from table \ref{tab:vertex} to the simulated data; the resulting cuts are shown in figure~\ref{simvertex}.

@@

We used a 4-dimensional binning to divide the large phase space of our two particle final 
state. However, to evaluate systematic errors associated with this correction, we compared the
method using two different sets of kinematic variables, as presented in 
table~\ref{tab:AcceptBinning} and \ref{tab:AcceptBins}. The first set of variables is the one actually used for the
analysis and the second set is only used to evaluate the systematic errors, so in the 
following, the formulas are given only for the first set. The total number of bins is 
constrained by the amount of generated data and must be maintained reasonably low in order 
to keep a good statistical precision and a small bin migration.

@@

Because of reconstruction errors or scattering on detector material, it 
is possible to associate a particle with the wrong target. To estimate this 
effect, we looked, in the experimental data, at the number of reconstructed 
events upstream and downstream of our targets, where nothing should be 
detected. Thus, we defined two test regions (see figure \ref{fig:targetleak}) 
to evaluate the associated contamination between the targets. The upstream 
region, named test region 1, is of the size of the solid target's selection 
cut to evaluate the contamination from the liquid target. The downstream region, 
named test region 2, covers the same size of the liquid target's selection cut 
to estimate the leak from the solid target. The distance between detection and 
test regions is identical to the distance between the two targets' detection 
regions.

@@

The point-to-point errors include contributions from particle 
misidentifications and acceptance effects. The normalization error, which is 
independent of kinematic variables, are due to acceptance effects, target 
misidentification and the isospin correction. The various constributions 
to the systematic normalization error are summarized in table~\ref{tab:sysid}.

### 2017-01-13 chap6.tex
Finished English corrections

The attenuation, presented in the figure \ref{fig:RA}, is neither linear as a 
function of $A^{1/3}$ nor as a function of $A^{2/3}$. The HERMES data had
already showed some indication 
of this feature~\cite{Airapetian:2007vu,Airapetian:2009jy}, but here the 
non-linearity, which has important implication on models, is clear. Indeed, 
it seems difficult to conciliate the prehadron absorption models with this result. Prehadrons 
are expected to have their cross sections increasing with time and, therefore, 
distance. On the parton energy loss side, the BDMPS calculation gives a parton 
energy loss proportional to $L^2 \propto A^{2/3}$, indicating that we are not
in the region where this result applies. Morover, if the production 
time\footnote{As defined in figure~\ref{fig:hadro}} occurs 
inside the nuclei, the behavior of the multiplicity ratio can be more complicated
as different processes intervene.

@@

The result of the lead to carbon multiplicity ratio, figure \ref{fig:RPt} 
(right), has another interesting feature: the $\pi^+$s have a stronger Cronin 
effect. However, the signification of this observation is not clear. This could be 
a contribution from target fragmentation, as we might expect more positive 
particles would originate from nuclei than negative ones. This could also come from other sources such as the isospin effect or a factorization breaking at high 
\ptp. However, no test of these features exists in our kinematical range\footnote{Results of \cite{Asaturyan:2011mq} covered only \pt$<0.2$ GeV$^2$/c$^2$.}. Last but not least, it could be correlated with the hadronic rank and reflect some hadronization properties. Indeed, as we probe mostly valence quarks at our energies, $\pi^+$s are more likely to be the leading hadrons compared to $\pi^-$s.

@@

The HERMES collaboration clearly observed a rise of the multiplicity ratio with $\nu$ from 4 till 22~GeV. However, at first sight to our results (figure \ref{fig:RNu} (left)), no dependence is observed, besides a slight increase in lead.
However, the Fermi motion could cancel most of this hadronization effect, and what left will be washed out with the acceptance correction systematic uncertainty. The normalized multiplicity ratio to carbon, however, offers a cleaner result with a slope consistent with what was observed on a much larger range of HERMES data.

@@

The behavior of hadronization as a function of $Q^2$ is an important issue, that 
has direct implications on our understanding of nuclear matter properties in 
QCD. HERMES results, which covered $1<Q^2<10$~GeV$^2$/c$^2$, gave a hint of an 
increase of the multiplicity ratio with $Q^2$. Our result, in figure 
\ref{fig:RQ2}, did not indicate any significant $Q^2$ dependence, hence no conclusions can be made from our measurement.

@@

The $A$ dependence of the transverse momentum broadening, presented in 
figure~\ref{fig:PA}, is a crucial result. The combination of our large 
statistics with a wide $A$ coverage gave an outright indication of $A^{1/3}$ 
dependence of \dpt. This \dptp~ effect is found much smaller compared to 
HERMES measurment\cite{Airapetian:2009jy}, which is consistent with
 theoretical models predicting larger effects at larger energies. However, 
calculations of the parton energy loss from BDMPS \cite{Baier:1996sk} 
correlated \pt with quadratically the nuclei radii. A feature that is not 
observed in our result. Although, by looking to the low energy of our CLAS 
experiment, one might argue the validity of a direct comparison with BDMPS 
calculations. Nonetheless, this result shows an unexpected pattern that 
remains to be explained. One possible interpretation is the occurrence of 
the production time inside the nuclei, as we proposed for the $A^{1/3}$ 
dependence of multiplicity ratios. In this case, the colored parton would 
not interact with the whole nuclei, hence limiting the nuclear effect.

@@

Finally, the $Q^2$ dependence of \dpt is an important result for the BDMPS based 
calculation from \cite{Domdey:2008aq}. They have expected a raise of \dpt with $Q^2$, 
which is not observed in figure \ref{fig:PQ2}. Using a tighter $\nu$ bin and a 
carbon normalization should isolate this effect better, but figure \ref{fig:PQ2-detailed}
gives a similar result. In conclusion, within error bars, no effect is observed for 
\dpt as a function of $Q^2$. However, we would like to have a quantitative theoretical 
input here, as it is not clear if we have the needed resolution to observe the effect 
expected within BDMPS based models. We are going to work with the authors of the original 
paper to add some projections on our figures.

>>>>

The attenuation, presented in the figure \ref{fig:RA}, is neither linear as a 
function of $A^{1/3}$ nor as a function of $A^{2/3}$. The HERMES data had
already showed some indication 
of this feature~\cite{Airapetian:2007vu,Airapetian:2009jy}, but here the 
non-linearity, which has important implication on models, is clear. Indeed, 
it seems difficult to reconcile the prehadron absorption models with this result. Prehadrons 
are expected to have their cross sections increasing with time and, therefore, 
distance. On the parton energy loss side, the BDMPS calculation gives a parton 
energy loss proportional to $L^2 \propto A^{2/3}$, indicating that we are not
in the region where this result applies. Morover, if the production 
time\footnote{As defined in figure~\ref{fig:hadro}} occurs 
inside the nuclei, the behavior of the multiplicity ratio can be more complicated
as different processes intervene.

@@

The result of the lead to carbon multiplicity ratio, figure \ref{fig:RPt} 
(right), has another interesting feature: the $\pi^+$s have a stronger Cronin 
effect. However, the significance of this observation is not clear. This could be 
a contribution from target fragmentation, as we might expect more positive 
particles would originate from nuclei than negative ones. This could also come from other sources such as the isospin effect or a factorization breaking at high 
\ptp. However, no test of these features exists in our kinematical range\footnote{Results of \cite{Asaturyan:2011mq} covered only \pt$<0.2$ GeV$^2$/c$^2$.}. Last but not least, it could be correlated with the hadronic rank and reflect some hadronization properties. Indeed, as we probe mostly valence quarks at our energies, $\pi^+$s are more likely to be the leading hadrons compared to $\pi^-$s.

@@

The HERMES collaboration clearly observed a rise of the multiplicity ratio 
with $\nu$ from 6 to 22~GeV. However, our results (figure \ref{fig:RNu} (left))
show no dependence, besides a slight increase in lead.
However, the Fermi motion could cancel most of this hadronization effect, and 
what is left will be washed out with the acceptance correction systematic 
uncertainty. The normalized multiplicity ratio to carbon offers a 
cleaner result, that is not affected by Fermi motion or acceptance. For this
observable, we measure a slope consistent with what was observed by HERMES.

@@

The behavior of hadronization as a function of $Q^2$ is an important issue, that 
has direct implications on our understanding of nuclear matter properties in 
QCD. HERMES results, which covered $1<Q^2<10$~GeV$^2$/c$^2$, gave a hint of an 
increase of the multiplicity ratio with $Q^2$. Our result, in figure 
\ref{fig:RQ2}, does not indicate any $Q^2$ dependence, hence our 
conclusion is that the multiplicity ratio has no significant $Q^2$ dependence.

@@

The $A$ dependence of the transverse momentum broadening, presented in 
figure~\ref{fig:PA}, is a crucial result. The combination of our large 
statistics with a wide $A$ coverage gives an outright indication of $A^{1/3}$ 
dependence of \dpt. This \dpt effect is found to be much smaller than that seen 
by HERMES\cite{Airapetian:2009jy}, which is consistent with
 theoretical models predicting larger effects at larger energies. However, 
calculations of the parton energy loss from BDMPS \cite{Baier:1996sk} 
correlated \pt with the square of the nuclear radius. A feature that is not 
observed in our result. Although, by looking to the low energy of our CLAS 
experiment, one might argue the validity of a direct comparison with BDMPS 
calculations. Nonetheless, this result shows an unexpected pattern that 
remains to be explained. One possible interpretation is the occurrence of 
the production time inside the nuclei, as we proposed for the $A^{1/3}$ 
dependence of multiplicity ratios. In this case, the colored parton does 
not interact with the whole nucleus which limits the nuclear effect.

@@

Finally, the $Q^2$ dependence of \dpt is an important result for the BDMPS based 
calculation from \cite{Domdey:2008aq}. They predict a raise of \dpt with $Q^2$, 
which is not observed in figure \ref{fig:PQ2}. Using a tighter $\nu$ bin and a 
carbon normalization should isolate this effect better, but figure \ref{fig:PQ2-detailed}
gives a similar result. In conclusion, within error bars, no effect is observed for 
\dpt as a function of $Q^2$. However, we would like to have a quantitative theoretical 
input here, as it is not clear if we have the needed resolution to observe the effect 
expected within BDMPS based models. We are going to work with the authors of the original 
paper to add some projections on our figures.
